Bird beak's elaboration and innovation

### 2023-04-27 Analysis/07-Orthogonality_and_statistics.Rmd
Update 07-Orthogonality_and_statistics.Rmd

```{r, eval = TRUE, echo = FALSE}
source("../Functions/stats.utilities.R")

@@

>>>>

```{r, eval = FALSE, echo = FALSE}
source("../Functions/stats.utilities.R")

@@

## Random skewers

```{r}
## Random skewers function wrapper for dispRity
skewers.fun <- function(matrix, matrix2) {
    return(phytools::skewers(matrix, matrix2)$r)
}
skewers.fun.p <- function(matrix, matrix2) {
    return(phytools::skewers(matrix, matrix2)$p)
}
p.star <- function(x) {
    if(x > 0.1) {
        return("")
    } else {
        if(x > 0.05) {
            return(".")
        } else {
            if(x > 0.01) {
                return("*")
            } else {
                if(x > 0.001) {
                    return("**")
                } else {
                    return("***")
                }
            }
        }
    }
}

## Getting the same order as the orthogonality data
ortho_order <- c(
    "Columbimorphae",
    "Galloanserae",
    "Mirandornithes",
    "Telluraves",
    "Strisores",
    "Paleognathae",
    "Aequornithes",
    "Otidimorphae",
    "Galliformes",
    "Accipitriformes",
    "Strigiformes",
    "Musophagiformes",
    "Trogoniformes",
    "Columbimorphae",
    "Caprimulgiformes",
    "Bucerotiformes",
    "Falconiformes",
    "Pterocliformes",
    "Sphenisciformes",
    "Podicipediformes",
    "Ciconiiformes",
    "Tinamiformes",
    "Suliformes",
    "Anseriformes",
    "Apodiformes",
    "Otidiformes",
    "Cuculiformes",
    "Psittaciformes",
    "Gruiformes",
    "Pelecaniformes",
    "Procellariiformes",
    "Piciformes",
    "Passeriformes",
    "Coraciiformes",
    "Charadriiformes")

## Getting the group pairs
group_pairs <- sapply(name.subsets(all_birds_shapespace)[-36], function(x) c(x, "phylogeny"), simplify = FALSE)

## Calculating the random skewers for each group
skewers_r <- dispRity(all_birds_shapespace, metric = as.covar(skewers.fun), between.groups = group_pairs, verbose = TRUE)
skewers_p <- dispRity(all_birds_shapespace, metric = as.covar(skewers.fun.p), between.groups = group_pairs, verbose = TRUE)

## Extract disparity for fancy plotting
skewers_distrib <- lapply(get.disparity(skewers_r, concatenate = FALSE), quantile, probs = c(0.5, 0.025, 0.25, 0.75, 0.975))
skewers_median_p <- lapply(get.disparity(skewers_p, concatenate = FALSE), median)
plot_names <- unlist(strsplit(names(skewers_distrib), split = ":phylogeny"))
names(skewers_distrib) <- names(skewers_median_p) <- plot_names
skewers_distrib <- skewers_distrib[ortho_order]
skewers_median_p <- skewers_median_p[ortho_order]
plot_names <- names(skewers_median_p)

col <- rev(c(rep(gg.color.hue(4)[1], 8), rep(gg.color.hue(4)[2], 27)))

par(mar = c(5, 17, 4, 2))
plot(NULL, xlim = c(0, 1), ylim = c(1,length(skewers_distrib)), xlab = "Skewers correlation", ylab = "", yaxt = "n")
abline(v = 0.5, lty = 2, col = "grey", lwd = 0.5)
axis(2, at = length(skewers_distrib):1, labels = plot_names, las = 2)
## Skewers
for(i in length(skewers_distrib):1) {
    lines(x = skewers_distrib[i][[1]][c(2, 5)], y = rep(i, 2), lty = 2, lwd = 1, col = col[i])
    lines(x = skewers_distrib[i][[1]][c(3, 4)], y = rep(i, 2), lty = 1, lwd = 2, col = col[i])
    points(x = skewers_distrib[i][[1]][1], y = i, pch = 19, col = col[i])
    text(x = skewers_distrib[i][[1]][1]+0.03, y = i+0.2, label = p.star(skewers_median_p[[i]]), col = col[i])
}
